--- 
title: "As a Statistician"
author: "Zehui Bai"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This book involves different statistical principles and methods, including the application of SAS and R, and aims to accumulate personal statistical knowledge."
---

# Introduction

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```



You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].


**The main content contains the following modules**:

* Data Management

    + Chap 01 Introduction
    + Chap 02 Input and Format
    + Chap 03 DM in SAS
    + Chap 04 DM in R
    + Chap 05 SQL
  
* Data Visualization  
    + Chap 07 ggplot2
    + Chap 08 Proc Sqplot
    
* Data Analysis

    + Chap 11 Descriptive Statistics in SAS
    + Chap 12 Descriptive Statistics in R
    
    + Chap 13 Distributions
    + Chap 14 Parametric Test
    + Chap 15 Multiple Comparison
    + Chap 16 CI and Sample Size Calculation
    + Chap 17 Non-Parametric Test
    
    + Chap 18 Correlation and Regression
    + Chap 19 Advanced Linear Regression
    
    + Chap 20 Logistic Regression 
    + Chap 21 Advanced Logistic Regression 
    
    + Chap 22 Survival Analysis
    + Chap 23 Advanced Survival Analysis
    
    + Chap 24 Count Data Regression
    + Chap 25 Proportion Response Regression
    
    + Chap 26 Mixed Model
    + Chap 27 Generalized Linear (Mixed) Model
    + Chap 28 Generalized Estimating Equation
    
    + Chap 29 Time Series Analysis
    + Chap 30 Meta Analysis
    
    + Chap 31 Group (Adaptive) Sequential Design
    + Chap 32 Clinic Study Design
    
    + Chap 33 Propensity Score
    
    + Chap 34 Missing Data
    + Chap 35 
    
* Machine learning and Deep learning

    + Chap 38 Bayesian Theory     
    + Chap 39 Bayesian Regression
    
    + Chap 43 Regularization: Penalized Regression
    + Chap 44 
    
    + Chap 45 Classification Analysis 
    + Chap 46 Random Forest
    + Chap 47 Gradient Descent and Gradient Boosting
    + Chap 48 Cubist Model
    + Chap 49 Support Vector Machine (SVM)
    + Chap 50 K-Nearest Neighbors (KNN)
    + Chap 51 Linear Discriminant Analysis (LDA)
    + Chap 52 Multivariate Adaptive Regression Spline (MARS)
    + Chap 53 Clustering Analysis
    
    + Chap 56 Singular Value Decomposition (SVD)
    + Chap 57 Principal Component Analysis (PCA) 
    
    + Chap 60 Neural Networks
    
    
